Maximizing Ink in Symmetric Partial Edge Drawings of $k$-plane Graphs

### 2018-01-09 eurocg/paper.tex
minor edits

\affil[1]{Algorithms and Complexity Group, TU Wien, Vienna, Austria\\ \texttt{[fklute|noellenburg]@ac.tuwien.ac.at}}
%\affil[2]{Algorithms and Complexity Group, TU Wien, Vienna, Austria} %\texttt{fklute@ac.tuwien.ac.at}}
%\affil[3]{Algorithms and Complexity Group, TU Wien, Vienna, Austria \texttt{}}
%\affil[4]{Algorithms and Complexity Group, TU Wien, Vienna, Austria} %\texttt{noellenburg@ac.tuwien.ac.at}}
%\affil[5]{Algorithms and Complexity Group, TU Wien, Vienna, Austria \texttt{}}
%mandatory. First: Use abbreviated first/middle names. 
%Second (only in severe cases): Use first author plus 'et. al.'
\authorrunning{M. Höller, F. Klute, S. Nickel, M. N\"ollenburg, and B. Schreiber}

@@

%Let $ C = (V,E) $ be a cactus graph and moreover the intersection graph of a graph $ G $ and straight-line drawing $ \Gamma $. Additionally let $ \mathcal T = (\mathcal V, \mathcal E) $ be the block-cut tree of $ C $ 
We further define for each node $ u \in \mathcal V $ two indices $ i_p(u) \le i_q(u) $. %, w.l.o.g $ i_p > i_q $. 
For an articulation node $u$, $i_p$ and $i_q$ are the indices corresponding to the stub lengths induced by the intersection points of $s(a(u))$ and its neighbors in $C$ among the vertex set of $p(u)$.
For a block node $u$, $i_p$ and $i_q$ are the indices corresponding to the intersection points of the parent's segment $s(a(p(u)))$ with its neighbors in $V(u)$.
%These are exactly the indices of the stub-lengths leading to an intersection between two cycles of segments in $ \Gamma $. 
Similar to Section~\ref{sec:tree} we now define three values for each articulation node $ u \in \mathcal V $ with degree $ \delta $:
\begin{align*}
	\sollong(u) = \max\{T_0(u),\dots,T_\delta(u)\}\\
	\solmid(u) = \max\{T_1(u),\dots,T_{i_q}(u)\}\\
	\solshort(u) = \max\{T_1(u),\dots,T_{i_p}(u)\}
\end{align*}
Further we define the recurrence $T_i(u)$ for all articulation nodes $u \in \mathcal V$:

\begin{align}
\label{rec:cactus_articulation}
T_i(u) = \ell_i(a(u)) + \sum_{v \in c(u)}
\begin{cases}
%\solshort(v) & \text{if } \ell_i(A(u)) \leq a \\
%\solmid(v) & \text{if } a < \ell_i(A(u)) \leq b \\
%\sollong(v) & \text{else}
%\end{cases}.
%\end{align}
%For a block vertex $ u \in \mathcal V $ we have to cut open the cycle it represents (compare \cite{bcgkmn-pped-17}). We do this at $ A(p(u)) $. The resulting graph is a path between the vertices in $ \textit{cycle}(u) $ with attached articulation vertices. This is especially a tree and we use the aglorithm from Section~\ref{sec:tree}, maximizing over three values of the articulation vertices though,  to compute the $ T_i(u) $.
%
%If the values computed for the block vertices is correct, we can follow the correctness of the algorithm as in Section~\ref{sec:tree}. Let $ u \in \mathcal V $ be a block vertex then by \cite{bcgkmn-pped-17} we know the optimal solution for \maxsped just on $ C[\textit{cycle}(u)] $ is correctly computed by the above procedure since the algorithm from Section~\ref{sec:tree} correctly handles the created path. Now if $ u $ has articulation vertices attached to it the above computes the correct value since this is simply a tree. Further there is no $ v \in c(u) $ which is not an articulation vertex, hence the maximum ink value for $ u $ is correctly computed if we know the values for the children in $ c(u) $.
%
%For the runtime we can again argue the same as in Section~\ref{sec:tree}, but this time we have to also check cycles of length $ n_c $ which leads to a runtime of $ O(n_ck^2) $ for a block vertex $ u \in \mathcal V $ with $ n_c = |\textit{cycle}(u)| $. In case $ u $ is an articulation vertex we can solve it in $ O(k) $. As before the optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $ can be found by backtracking from the root. Additionally the intersection graph can be computed in $ O(m\log m) $ where $ m $ is the number of edges in the original graph $ G $ since $ |E| = O(m) $ in a cactus and we obtain the following theorem.
T_{i_p(v)}(v) & \text{if } i < i_p(v)\\ %\ell_i(a(u)) \leq a \\
T_{i_q(v)}(v) & \text{if } i_p(v) < i \leq i_q(v) \\
T_{0}(v) & \text{otherwise}
% \solshort(v) & \text{if } i < i_p(v)\\ %\ell_i(a(u)) \leq a \\
% \solmid(v) & \text{if } i_p(v) < i \leq i_q(v) \\
% \sollong(v) & \text{otherwise}
\end{cases}.
\end{align}
For a block node $ u \in \mathcal V $ we cut open the cycle it represents at vertex $a(p(u))$ (refer to Bruckdorfer et al.~\cite{bcgkmn-pped-17} for details) and consider the three relevant stub lengths of the parent's segment $s(a(p(u)))$ induced by $i_p(u)$ and $i_q(u)$ and the whole segment. 
The resulting subgraph is a path on $V(u)$ with attached children that are articulation vertices in $\mathcal T$ so we can apply our algorithm of Section~\ref{sec:tree} with minor modifications.
Once all values in the subgraph of block node $u$ are computed, we can derive the values $T_i(u)$ for the block node $u$.

>>>>

\affil[1]{Algorithms and Complexity Group, TU Wien, Vienna, Austria\\ \texttt{[fklute|noellenburg]@ac.tuwien.ac.at, soeren.nickel@tuwien.ac.at}}
%\affil[2]{Algorithms and Complexity Group, TU Wien, Vienna, Austria} %\texttt{fklute@ac.tuwien.ac.at}}
%\affil[3]{Algorithms and Complexity Group, TU Wien, Vienna, Austria \texttt{}}
%\affil[4]{Algorithms and Complexity Group, TU Wien, Vienna, Austria} %\texttt{noellenburg@ac.tuwien.ac.at}}
%\affil[5]{Algorithms and Complexity Group, TU Wien, Vienna, Austria \texttt{}}
%mandatory. First: Use abbreviated first/middle names. 
%Second (only in severe cases): Use first author plus 'et. al.'
\authorrunning{M. Höller, F. Klute, S. Nickel, M. N\"ollenburg, and B. Schreiber}

@@

%Let $ C = (V,E) $ be a cactus graph and moreover the intersection graph of a graph $ G $ and straight-line drawing $ \Gamma $. Additionally let $ \mathcal T = (\mathcal V, \mathcal E) $ be the block-cut tree of $ C $ 
We further define for each node $ u \in \mathcal V $ two indices $ i_p(u) \le i_q(u) $. %, w.l.o.g $ i_p > i_q $. 
For an articulation node $u$, $i_p$ and $i_q$ are the indices corresponding to the stub lengths induced by the intersection points of $s(a(u))$ and its neighbors in $C$ among the vertex set of $p(u)$.
For a block node $u$, $i_p$ and $i_q$ are the indices corresponding to the intersection points of the parent's segment $s(a(p(u)))$ with its neighbors in $V(u)$.
%These are exactly the indices of the stub-lengths leading to an intersection between two cycles of segments in $ \Gamma $. 
Similar to Section~\ref{sec:tree} we now define three values for each articulation node $ u \in \mathcal V $ with degree $ \delta $:
\begin{align*}
	\sollong(u) = \max\{T_0(u),\dots,T_\delta(u)\}\\
	\solmid(u) = \max\{T_1(u),\dots,T_{i_q}(u)\}\\
	\solshort(u) = \max\{T_1(u),\dots,T_{i_p}(u)\}.
\end{align*}
Further we define the recurrence $T_i(u)$ for all articulation nodes $u \in \mathcal V$:
\label{rec:cactus_articulation}
T_i(u) = \ell_i(a(u)) + \sum_{v \in c(u)}
%\solshort(v) & \text{if } \ell_i(A(u)) \leq a \\
%\solmid(v) & \text{if } a < \ell_i(A(u)) \leq b \\
%\sollong(v) & \text{else}
%\end{cases}.
%\end{align}
%For a block vertex $ u \in \mathcal V $ we have to cut open the cycle it represents (compare \cite{bcgkmn-pped-17}). We do this at $ A(p(u)) $. The resulting graph is a path between the vertices in $ \textit{cycle}(u) $ with attached articulation vertices. This is especially a tree and we use the aglorithm from Section~\ref{sec:tree}, maximizing over three values of the articulation vertices though,  to compute the $ T_i(u) $.
%If the values computed for the block vertices is correct, we can follow the correctness of the algorithm as in Section~\ref{sec:tree}. Let $ u \in \mathcal V $ be a block vertex then by \cite{bcgkmn-pped-17} we know the optimal solution for \maxsped just on $ C[\textit{cycle}(u)] $ is correctly computed by the above procedure since the algorithm from Section~\ref{sec:tree} correctly handles the created path. Now if $ u $ has articulation vertices attached to it the above computes the correct value since this is simply a tree. Further there is no $ v \in c(u) $ which is not an articulation vertex, hence the maximum ink value for $ u $ is correctly computed if we know the values for the children in $ c(u) $.
%For the runtime we can again argue the same as in Section~\ref{sec:tree}, but this time we have to also check cycles of length $ n_c $ which leads to a runtime of $ O(n_ck^2) $ for a block vertex $ u \in \mathcal V $ with $ n_c = |\textit{cycle}(u)| $. In case $ u $ is an articulation vertex we can solve it in $ O(k) $. As before the optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $ can be found by backtracking from the root. Additionally the intersection graph can be computed in $ O(m\log m) $ where $ m $ is the number of edges in the original graph $ G $ since $ |E| = O(m) $ in a cactus and we obtain the following theorem.
T_{i_p(v)}(v) & \text{if } 1 \le i < i_p(v)\\ %\ell_i(a(u)) \leq a \\
T_{i_q(v)}(v) & \text{if } i_p(v) < i \leq i_q(v) \\
T_{0}(v) & \text{otherwise}
% \solshort(v) & \text{if } i < i_p(v)\\ %\ell_i(a(u)) \leq a \\
% \solmid(v) & \text{if } i_p(v) < i \leq i_q(v) \\
% \sollong(v) & \text{otherwise}
For a block node $ u \in \mathcal V $ we cut open the cycle it represents at vertex $a(p(u))$ (refer to Bruckdorfer et al.~\cite{bcgkmn-pped-17} for details) and consider the three relevant stub lengths of the parent's segment $s(a(p(u)))$ induced by $i_p(u)$ and $i_q(u)$ and the whole segment. 
The resulting subgraph is a path on $V(u)$ with attached children that are articulation vertices in $\mathcal T$ so we can apply our algorithm of Section~\ref{sec:tree} with minor modifications.
Once all values in the subgraph of block node $u$ are computed, we can derive the values $T_i(u)$ for the block node $u$.
